Galaxy Zoo: Morphological Classifications for 150,000 Galaxies in HST Legacy Imaging

### 2016-05-04 writeup/gz_hubble_data.tex
added ferengi stats and updated figures in appendices

\subsection{Challenges of debiasing questions beyond ``smooth or features''}

As with the HST images, each FERENGI subject has a varying number of users answering the various questions in the hierarchical decision tree. Every user answers the first question, ``Is the galaxy smooth and rounded, with no sign of a disk?''; as such the vote fractions $\rm p_{smooth}, p_{features}$, and $\rm p_{artifact}$ are all computed with the minimum statistical error for any question, with roughly 40 total answers (see Section~\ref{sec:interface}). The number of users to answer any subsequent question, however, is always equal to or less than the number to answer the preceding question. The average number of responses per task for fourth- or fifth-tier questions such as spiral arm structure (Tasks 12--14) is only $4\pm4$ for the FERENGI sample; while this is distribution is strongly bimodal (reflecting the true morphologies of selected galaxies), the very low absolute numbers of votes introduce very high variance when attempting to calculate a statistical correction.

In the FERENGI data, these numbers severely limit the amount of information we can extract for the higher-tier questions. The debiasing technique used (Section~\ref{ssec:zeta}) require that at least 10~users answer each question for a galaxy image at $z=0.3$ \emph{and} its image at higher redshift. This requirement is by default met by all galaxies when analyzing the first question, since it is asked of all users. This requirement is often \emph{not} met for questions beyond Task~01; on average 60.0\% of the galaxies do not have sufficient data to measure a correction, as compared to 2.0\% acheived for Task~01 (see Table~\ref{tbl:ferengi_corrections}). Of the remaining FERENGI galaxies which did meet this requirement, polynomials were fit to the data in each redshift-surface brightness bin, as was done for Task~01 (described above). The goodness-of-fit was evaluated in each bin, for each Task, using the value of the sum of the square residuals (SSR). The mean SSR for all Tasks was $\mu_{\rm SSR}$ = 0.46, $\sigma$ = 0.36, which was not considered sufficiently low in any Task for the fits to accurately model the data. An example of the debiasing technique applied to $p_{\rm bar}$ is shown in Section \ref{sec:ferengi_bar}; visual inspection of Figure~\ref{fig:p_vs_p_bar} confirms the available data in each bin is scattered to a degree at which no correlation may be confidently obtained. For these reasons, debiased vote fractions are only offered for Task~01 (smooth/features).  
\begin{table*}
\caption{Correctable fractions for the top-level task in GZH, split by survey.}\label{tbl:hubble_debiasable}
\begin{tabular}{lcrrrrrrr}
\hline\hline
                                   & Correction type & AEGIS   & COSMOS & GEMS & GOODS-N & GOODS-S    & SDSS    & Total \\
\hline
Correctable                        & 0               & 1,654   & 15,170 & 1,837 & 993    & 835     	& 0       & 20,489\\
Lower-limit                        & 1               & 1,917   & 26,113 & 2,423 & 1,385  & 1,282   	& 0       & 33,120\\
No Correction Needed ($z \le 0.3$) & 2               & 955     & 11,926 & 1,175 & 415    & 400     	& 37,545  & 52,416\\ 
NEI                                & 3               & 2,847   & 34,511 & 3,308 & 2,535  & 2,523   	& 0       & 45,724\\
No Redshift Information            & 4               & 1,134   & 5,088  & 561   & 687    & 102   		& 14,316  & 21,888\\
Total                              &                 & 8,507   & 92,808 & 9,304 & 6,015  & 5,142   	& 51,861  & 173,637\\
\hline\hline
\end{tabular}
\end{table*}

@@

\begin{figure}
\begin{center}
\includegraphics[width=0.50\textwidth]{figures/full_shallow_p_plot.pdf}
\caption{shallowfull}
\label{fig:shallow_vs_full}
\end{center}
\end{figure}

@@

\subsection{FERENGI analysis of pbar}
\label{sec:ferengi_bar}
\begin{figure*}
\centering
\subfigure{[a]\label{a}\includegraphics[width=0.5\textwidth]{figures/p_vs_p_SB_redshift_bar.pdf}}
\subfigure{[b]\label{b}\includegraphics[width=0.4\textwidth]{figures/z1_mu20_subplot1_bar.pdf}}
\\
\subfigure{[c]\label{c}\includegraphics[width=0.5\textwidth]{figures/orangebars_bar.pdf}}
\subfigure{[d]\label{d}\includegraphics[width=0.4\textwidth]{figures/z1_mu20_subplot2_bar.pdf}}
\caption{Same as Figure~\ref{fig:p_vs_p}, but with the bar question.}
\label{fig:p_vs_p_bar}
\end{figure*}

\begin{table}ー
\caption{Distribution of \ferengi{} images analysed in Figure~\ref{fig:p_vs_p_bar}. Correctable images had a single-valued relationship between their measured \pbar{} values at high and low redshifts (white regions in Figure~\ref{fig:p_vs_p_bar}). Galaxies with a lower-limit on \pbar{} had a non single-valued relationship (blue regions). NEI images had undetermined relationships due to a lack of data ($N<5$) in their corresponding $z$-$\mu$ bins (gray regions). Only 17\% (maximum) of FERENGI galaxies in the sample were considered ``correctable'', which is not sufficient to compute a $\zeta$ function applicable to the Hubble data.   \label{tbl:ferengi_bar_corrections}}
\begin{tabular}{lrr}
\hline \hline
				                   & N       & \% \\
\hline 
Correctable                        & 664   & 17\% \\
Lower-limi                         & 483   & 12\% \\
NEI                                & 2,803     & 71\%\\
Total                              & 3,950   & 100\% \\
\hline \hline
\end{tabular}
\end{table}

>>>>

\subsection{Challenges of debiasing questions beyond ``smooth or features''}
\label{ssec:higher_order_tasks}
As with the HST images, each FERENGI subject has a varying number of users answering the various questions in the hierarchical decision tree. Every user answers the first question, ``Is the galaxy smooth and rounded, with no sign of a disk?''; as such the vote fractions $\rm p_{smooth}, p_{features}$, and $\rm p_{artifact}$ are all computed with the minimum statistical error for any question, with roughly 40 total answers (see Section~\ref{sec:interface}). The number of users to answer any subsequent question, however, is always equal to or less than the number to answer the preceding question. The average number of responses per task for fourth- or fifth-tier questions such as spiral arm structure (Tasks 12--14) is only $4\pm4$ for the FERENGI sample; while this is distribution is strongly bimodal (reflecting the true morphologies of selected galaxies), the very low absolute numbers of votes introduce very high variance when attempting to calculate a statistical correction.

In the FERENGI data, these numbers severely limit the amount of information we can extract for the higher-tier questions. The debiasing technique used (Section~\ref{ssec:zeta}) require that at least 10~users answer each question for a galaxy image at $z=0.3$ \emph{and} its image at higher redshift. This requirement is by default met by all galaxies when analyzing the first question, since it is asked of all users. This requirement is often \emph{not} met for questions beyond Task~01; on average 60.0\% $\pm$ 24.0\% of the galaxies do not have sufficient data to measure a correction, as compared to 2.0\% acheived for Task~01 (see Table~\ref{tbl:ferengi_corrections}). Of the remaining FERENGI galaxies which did meet this requirement, polynomials were fit to the data in each redshift-surface brightness bin, as was done for Task~01 (described above). The goodness-of-fit was evaluated in each bin, for each Task, using a normalized $\chi^2$ metric. The mean $\chi^2$ for all higher-order tasks was $\mu_{\chi^2} = 0.10 \pm 0.04$, which were significantly greater than that achieved for Task 01, $\chi^2 = 0.04$. From this statistic it was determined that the data could not be modeled with sufficient confidence for the higher order tasks. An example of the debiasing technique applied to $p_{\rm bar}$ is shown in Section \ref{sec:ferengi_bar}; visual inspection of Figure~\ref{fig:p_vs_p_bar} confirms the available data in each bin is scattered to a degree at which no correlation may be confidently obtained. For these reasons, debiased vote fractions are only offered for Task~01 (smooth/features).

\begin{table*}
\caption{Correctable fractions for the top-level task in GZH, split by survey.}\label{tbl:hubble_debiasable}
\begin{tabular}{lcrrrrrrr}
\hline\hline
                                   & Correction type & AEGIS   & COSMOS & GEMS & GOODS-N & GOODS-S    & SDSS    & Total \\
\hline
Correctable                        & 0               & 1,654   & 15,170 & 1,837 & 993    & 835     	& 0       & 20,489\\
Lower-limit                        & 1               & 1,917   & 26,113 & 2,423 & 1,385  & 1,282   	& 0       & 33,120\\
No Correction Needed ($z \le 0.3$) & 2               & 955     & 11,926 & 1,175 & 415    & 400     	& 37,545  & 52,416\\ 
NEI                                & 3               & 2,847   & 34,511 & 3,308 & 2,535  & 2,523   	& 0       & 45,724\\
No Redshift Information            & 4               & 1,134   & 5,088  & 561   & 687    & 102   		& 14,316  & 21,888\\
Total                              &                 & 8,507   & 92,808 & 9,304 & 6,015  & 5,142   	& 51,861  & 173,637\\
\hline\hline
\end{tabular}
\end{table*}

@@

\begin{figure*}
\begin{center}
\includegraphics[width=0.50\textwidth]{figures/full_shallow_p_plot.pdf}
\caption{Distribution of $p_{\rm features}$ for the 4,461 GOODS galaxies which had both shallow and full depth images morphologically classified in GZH. For most galaxies, the value of $p_{\rm features}$ is in strong agreement between both types of images. Examples of galaxies with both sharp changes in $p_{\rm features}$, and little to no change, are shown in Figures~\ref{fig:shallow_smooth}-\ref{fig:shallow_featured}.}
\label{fig:shallow_vs_full}
\end{center}
\end{figure*}

@@

\subsection{FERENGI analysis of pbar}
\label{sec:ferengi_bar}

To determine the rate at which the vote fraction for any Task decreases with redshift or surface brightness, data from simluated FERENGI images were modeled with functions in discrete redshift / surface brightness bins, described in detail in Section~\ref{ssec:zeta_results}. A cut of $N>10$ was placed on the number of votes to reduce the error in the vote fractions and ensure they were well-sampled before applying a fit to the data. This requirement was not met for sufficient data in Tasks higher than Task~01, which significantly reduced the goodness-of-fit for the model functions as compared with the features Task. Here we show an example of results obtained for the bar Task. In Figure~\ref{fig:p_vs_p_bar}, it is apparent that both the lack of points and the degree of scatter in each bin prevent proper functions to be fit to the data. Table~\ref{tbl:ferengi_bar} summarizes the correction-type classification of the FERENGI data for this Task; 71\% of the galaxies simulated did not meet requirements for well-sampled vote fractions, as compared with 2\% acheived for Task~01. These results were similar for all higher-order Tasks, as described in Section~\ref{ssec:higher_order_tasks}, and therefore corrections were not derived for these Tasks.

\begin{figure*}
\centering
\subfigure{[a]\label{a}\includegraphics[width=0.5\textwidth]{figures/p_vs_p_SB_redshift_bar.pdf}}
\subfigure{[b]\label{b}\includegraphics[width=0.4\textwidth]{figures/z1_mu20_subplot1_bar.pdf}}
\\
\subfigure{[c]\label{c}\includegraphics[width=0.5\textwidth]{figures/orangebars_bar.pdf}}
\subfigure{[d]\label{d}\includegraphics[width=0.4\textwidth]{figures/z1_mu20_subplot2_bar.pdf}}
\caption{Refer to Figure~\ref{fig:p_vs_p} for a comparison to the features question. Effects of redshift bias in 3,950 images in the \ferengi{} sample. [a]: Each point in a given redshift and surface brightness bin represents a unique galaxy. On the $y$-axis in each bin is the \pfeatures{} value of the image of that galaxy redshifted to the value corresponding to that redshift bin. On the $x$-axis is the \pfeatures{} value of the image of the same galaxy redshifted to $z=0.3$. The dashed black lines represent the best-fit polynomials to the data in each square. The solid black line represents \pfeaturesz=\pfeaturesrest. Regions in which there is a single-valued relationship between \pfeatures{} at high redshift and at $z=0.3$ are white; those in which there is not are blue, and those with not enough data ($N<5$) are gray. \textbf{The mean normalized $\chi^2$, $\mu_{\chi^2}=0.08$ over all bins confirmed the polynomials were not good fits to the data, and no correction was derived from these.} [b]: A larger version of the dark-outlined square in [a], containing \ferengi{} galaxies that have been artificially redshifted to $z=1.0$ and have surface brightnesses between $20.3 < \mu < 21.0$ $\rm (mag/arcsec^2)$. [c]: The same data as [a] is shown. Each $z,\mu$ bin is divided into 4 sub-bins to determine the range of intrinsic \pfeaturesrest{} for a given range of observed \pfeaturesz{} values. In each sub-bin, the orange bars represent the inner 80$^\mathrm{th}$ percentiles of the data, the boundaries of which are the lower and upper limits of the debiased values. [d]: The same data as [b], but highlighting the upper and lower limit regions.}
\label{fig:p_vs_p_bar}
\end{figure*}

\begin{table}ー
\label{tbl:ferengi_bar}
\caption{Distribution of \ferengi{} images analysed in Figure~\ref{fig:p_vs_p_bar}. Correctable images had a single-valued relationship between their measured \pbar{} values at high and low redshifts (white regions in Figure~\ref{fig:p_vs_p_bar}). Galaxies with a lower-limit on \pbar{} had a non single-valued relationship (blue regions). NEI images had undetermined relationships due to a lack of data ($N<5$) in their corresponding $z$-$\mu$ bins (gray regions). Only 17\% (maximum) of FERENGI galaxies in the sample were considered ``correctable'', which is not sufficient to compute a $\zeta$ function applicable to the Hubble data.   \label{tbl:ferengi_bar_corrections}}
\begin{tabular}{lrr}
\hline \hline
				                   & N       & \% \\
\hline 
Correctable                        & 664   & 17\% \\
Lower-limit                        & 483   & 12\% \\
NEI                                & 2,803     & 71\%\\
Total                              & 3,950   & 100\% \\
\hline \hline
\end{tabular}
\end{table}
